Paper

### 2017-11-27 Paper.Rmd
...

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')

get.legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

plot.theme = theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), strip.background =element_rect(fill="lightcyan2"),  
                   strip.text = element_text(size=14, face='bold'), 
                   axis.line=element_line(color='black'), panel.grid.major=element_line(color='grey90'),
                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing = unit(0.1, 'lines')  )

data = '~/Data/Ellie'
tile.w='5e+06_arms'
#tile.w=5e+06
#tile.w=1e7
#tile.w='arms'

@@

# Missing some of the samples as there were some normals that we removed from analysis
# for (pt in names(patient.data)) {
#   patient.data[[pt]]$info = patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
# }

# for (pt in names(validation.patient.data)) {
#   validation.patient.data[[pt]]$info = validation.patient.data[[pt]]$info[validation.patient.data[[pt]]$info$Samplename %in% colnames(validation.patient.data[[pt]]$seg.vals)[-(1:5)],]
# }

@@

```{r tile, echo=F, message=F, warning=F, include=F}
tileFile = paste('tile_patients_All_',as.character(tile.w),'.Rdata',sep='')
tile.files = grep(tileFile, analysis.files, value=T, fixed=T)
load(tile.files[grep('tile', tile.files)], verbose=T)

mergedDf = mergedDf[grep('X|Y',rownames(mergedDf), invert=T),]

mergedDfValidation = mergedDf[,validation.patient.info$Samplename]
dim(mergedDfValidation)

mergedDf = mergedDf[, intersect(colnames(mergedDf), patient.info$Samplename)]
dim(mergedDf)

cache.dir = paste(data, 'Analysis',sub('\\+', '', tile.w),   sep='/')
if (includeDemo)
  cache.dir = paste(cache.dir, '_with-demo', sep='')

`r pander(mergedDf[sort(sample(nrow(mergedDf), 10)), sample(ncol(mergedDf), 10) ], caption='Example rows/columns from the merged data')`

@@

# sort in label order
#subset(patient.info, Samplename %in% setdiff(colnames(mergedDf), names(labels))[-c(1:3)])
if (length(setdiff(colnames(mergedDf), names(labels))) > 3)
  warning("Labels vector is missing samples")

@@

cn.mtn.plot<-function(df, label, col=c("orange4","grey90","steelblue3"), type='seg') {
  p = ggplot(df, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', space='free_x', labeller=label)

  if (type == 'seg') {
    p = p + geom_segment(aes(x=start, xend=end, y=mean.value, yend=mean.value, color=cn), show.legend=T, size=1) + scale_color_manual(values=col) 
  }
  if (type == 'bar') {
    p = p + geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), show.legend=T) + scale_fill_manual(values=col) 
  }

  p + labs(x='', y='Mean adjusted segmentation value', title='All samples') +
    theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
          panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
          panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )
}

@@

locs = do.call(rbind.data.frame, lapply(rownames(mergedDf), function(x) unlist(strsplit(x,':|-'))))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))

@@

cn = function(x) {
  str = 'neutral'
  if(x <= -1) {
    str = 'loss'
  } else if (x >= 1) {
    str = 'gain'
  }
  return(str)
}
globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)

@@

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
model.performance<-function(p1se, coeffs ) {
  performance = do.call(rbind.data.frame, p1se)
  performance$alpha = rownames(performance)

  if (grepl('chr', performance$alpha[1])) {
    performance$alpha = factor(performance$alpha, levels=c(paste('chr',c(1:22), sep=''), setdiff(rownames(performance), paste('chr',c(1:22),sep=''))) )
  }

  ## check how often the feature(s) is selected at that lamda in each split. "stability selection"
  coef.stable = lapply( coeffs, function(cf) {
    sort(rowSums(cf[,-1]), decreasing=T)
  })
  coef.stable = coef.stable[names(which(sapply(coef.stable, length) > 0))]
  coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))

  performance = cbind(performance, do.call(rbind,  lapply(coef.stable, function(x) {
    cbind('n.Coef'=length(x), '25%'=length(which(x>=.25)), 
          '50%'=length(which(x>=.5)), '75%'=length(which(x>=.75 )), 
          '100%'=length(which(x==1)) )
  })))

  p = ggplot(performance, aes(alpha, mean, color=(alpha != 0.9))) + 
    geom_point(size=2) +
    geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
    geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), width=0.3, size=1) + 
    geom_text(aes(label=n.Coef), nudge_y=0.03, fontface='bold') +
    geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) + 
    scale_colour_manual(values=c('red', 'grey39')) + plot.theme + theme(legend.position = 'none') +
    labs(x='Elasticnet penalty value: ridge <-> lasso', y='Model classification at lambda-1se')

  cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=length(names(coef.stable)), 
                           dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
  for (i in names(coef.stable)) 
    cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
  return(list('performance'=performance, 'stable.coeffients'=coef.stable, 'cf.per.feature'=cfs,'plot'=p))
}

@@

write.table(ch, quote=F, sep='\t', row.names = F, col.names = T, file=paste(as.character(tile.w),'_features.txt'))
```

@@

oddPt = names(pg.samp)[which.max(performance.at.1se)]
subset(sum.patient.data, Hospital.Research.ID == oddPt)

pg.samp[[oddPt]]$Prediction

# Not the pt with the most samples or endos
sum.patient.data[which.max(sum.patient.data$total.samples),]

```{r, eval=F}
image(cov(dysplasia.df[names(which(labels == 1)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
roc.plot <- function(roc,title="") {
  df = cbind.data.frame('Specificity'=rev(roc$specificities), 'Sensitivity'=rev(roc$sensitivities))

  ggplot(df, aes(Specificity, Sensitivity)) + geom_line() + scale_x_reverse() + 
    geom_label(data=as.data.frame(t(round(pROC::coords(roc, "best"),2))), 
               aes(x=specificity, y=sensitivity, 
                   label=paste(' (FPR ', sprintf('%.1f',(1-specificity)*100), '%, TPR ', sprintf('%.1f',sensitivity*100),'%)\nAUC:', sprintf("%.2f",roc$auc*100), '%', sep='')), nudge_x=.25) +
  labs(title=title, x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
}

preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$Prediction))

@@

roc.plot(roc50,'Probability threshold = 0.5')

@@

roc.plot(roc, 'Probability threshold calculated by ROC')

@@

```{r p53roc, echo=F, fig.height=4, fig.width=4}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')

pROC::auc(p53roc)

@@

roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)')

pROC::auc(pathroc)

@@

roc.plot(hgdpathroc, 'Pathology (HGD,IMC)')

pROC::auc(hgdpathroc)
```

@@

```{r echo=F, fig.height=6, fig.width=8}

get.roc.stat<-function(roc) {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )

@@

dodge = position_dodge(width=0.9)
ggplot(melt(m, measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=0, vjust=-1 ), position=dodge, color='white')  +
  scale_fill_manual(values=c('purple3','dodgerblue3','firebrick3'), name='') + plot.theme + 
  theme(legend.position = 'bottom') + labs(x='', y='Percent', title='ROC Comparison')

@@

## Predictions laid out spatially

```{r, spacial, fig.width=8, fig.height=6}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
  )

@@

  tb = table(pt$Endoscopy.Year, pt$PID)
  tb[tb>0] = 1
  for (yr in names(which(rowSums(tb) > 1))) {
      tmp = pt[which(pt$Endoscopy.Year == yr),]
      tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))

@@

  pt$brks = cut(pt$months.before.final, include.lowest = T, ordered_result = T,
    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12),
    labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))

  tb = table(pt[,c('brks', 'PID')])
  for (i in 1:nrow(tb)) {
    pids = tb[i,]
    yrs = which(pids > 0)
    if (length(yrs) > 1) {
    for (j in 2:length(yrs)) 
      pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
    }
  }
  arrange(pt, -Endoscopy.Year, PID)
}

@@

cuts = seq(0,1,0.1)
#cuts = quantile(back$Prediction, probs=seq(0,1,1/5))

@@

pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

riskCols = c('red3','orange','green')
ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
  geom_vline(xintercept=pred.confidence$r2, linetype='longdash') +
  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=75, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
  scale_color_manual(values=c('grey','black')) + plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability')

@@

back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')

ggplot(back, aes(1:nrow(back), Prediction)) + geom_point(aes(color=Risk, shape=Status)) + labs(x='sample', y='model probability', title='Per-sample probabilities')

back = as.data.frame(back %>% group_by(Risk) %>% mutate( 'mean(P)'=mean(Prediction), 'se(P)'=sd(Prediction)/sqrt(length(Prediction)), 'sd(P)'=sd(Prediction) ))
```

@@

```{r jacknife, echo=F, warning=F, message=F}

coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
for (pt in names(pg.samp)) {
  lcf = loo.coefs[[pt]][,1]
  names(lcf) = rownames( loo.coefs[[pt]] )

  sharedcfs = intersect(names(lcf), ch$label)

  coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
}
coefs.per.pt[is.na(coefs.per.pt)] = 0

@@

ch$jack.se = apply(coefs.per.pt, 1, function(x) {
  jk = jackknife(x, var)
  jk$jack.se
})

ch$adj = ifelse(ch$coef<0,-1,1)
ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
  geom_errorbar(aes(ymin=(abs(coef)-jack.se)*adj, ymax=(abs(coef)+jack.se)*adj)) + coord_flip() +
  plot.theme + labs(x='', title='Jackknifed coefficient CI??')

@@

@@

```{r pathology, echo=F, message=F, warning=F, fig.height=8, fig.width=6}
model.pred<-function(pt, p53=F, pred = function(rows) length(which(rows$Prediction > cutoff))) {
  final.endo = subset(pt, Endoscopy.Year == max(Endoscopy.Year))
  final.endo = final.endo[nrow(final.endo),]

  h = subset(pt, Pathology %in% c('IMC','HGD'))
  l = subset(pt, Pathology == 'LGD' & PID != final.endo$PID)
  id = subset(pt, Pathology == 'ID' & PID != final.endo$PID)
  be = subset(pt, Pathology %in% c('BE') & PID != final.endo$PID)

  if (p53) {
    return(cbind('HGD'=pred(h), 'total.h'=nrow(h),#   nrow(subset(h, !is.na(p53.Status))), 
                  'LGD'=pred(l), 'total.l'=nrow(subset(l, !is.na(p53.Status))), 
                  'ID'=pred(id), 'total.id'=nrow(subset(id, !is.na(p53.Status))), 
                  'NDBE'=pred(be), 'total.be'=nrow(subset(be, !is.na(p53.Status)))))

  } else {
    return( cbind('HGD'=pred(h), 'total.h'=nrow(h), 
                  'LGD'=pred(l), 'total.l'=nrow(l), 
                  'ID'=pred(id), 'total.id'=nrow(id),
                  'NDBE'=pred(be), 'total.be'=nrow(be)) )
    }
}
collapse.pred.by.path<-function(pts, p53=F, predFUN = NULL) {
  if (!is.null(predFUN))
    by.path = do.call(rbind.data.frame, lapply(pts,model.pred,p53, predFUN) )
  else
    by.path = do.call(rbind.data.frame, lapply(pts,model.pred,p53) )

  sums.by.path = colSums(by.path)

  path = as.data.frame(matrix(sums.by.path, byrow=T, nrow=4, dimnames=list(c('HGD','LGD','ID','NDBE'),c('pred', 'total'))))
  path$prog.ratio = (path$pred/path$total)*100
  path$nonprog.ratio = 100-path$prog.ratio
  path$path = rownames(path)

  path$path = factor(path$path, levels=c('HGD','LGD','ID','NDBE')) 
  path$pred = ifelse(p53, 'p53','model')
  return(path)
}

global.path = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID])
global.pathNP = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID])

predP53 = function(rows) length(which(rows$p53.Status == 1 ))
p53byPath = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], T, predP53)
p53.melt = melt(p53byPath, measure.vars=c('prog.ratio', 'nonprog.ratio'), id.vars=c('path', 'total','pred'))
p53.melt$Status = ifelse(p53.melt$variable == 'prog.ratio', 'Progressor','Non-Progressor')

global.melt = melt(global.path, measure.vars=c('prog.ratio', 'nonprog.ratio'), id.vars=c('path', 'total','pred'))
p53.melt = melt(p53byPath, measure.vars=c('prog.ratio', 'nonprog.ratio'), id.vars=c('path', 'total','pred'))
np.melt = melt(global.pathNP, measure.vars=c('prog.ratio', 'nonprog.ratio'),id.vars=c('path','total','pred'))

global.melt$pred='model'
global.melt$status='Progressor'
np.melt$status='Non-Progressor'

#p53.melt$status='p53 IHC'
p53.melt = subset(p53.melt, variable == 'prog.ratio')
p53.melt$variable = paste('p53',p53.melt$variable,sep='')
p53.melt$status = 'Progressor'

path.m = rbind(global.melt, np.melt, p53.melt)

#m$status = factor(m$status, levels=c('Progressor', 'p53 IHC','Non-Progressor'), ordered = T)

path.m$pred = factor(path.m$pred)
path.m$final = ifelse(path.m$path == 'HGD', T, F)

ggplot(path.m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status) +
  geom_bar(stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('firebrick4','green4','slateblue4'), name='', labels=c('Pred. progressed', 'Pred. Not progressed','p53 IHC'))  +
  geom_text(aes(vjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')),
            position=position_stack(),color='white',show.legend=F) +
  geom_bar(data=p53.melt, aes(path, value, group=path), fill='slateblue4',stat='identity',position = position_dodge()) +
  geom_text(data=p53.melt, aes(path,value,vjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')), position=position_dodge(),color='white',  show.legend=F) +
  labs(title='Samples predicted by pathology', x='Pathology', y='Percent', subtitle='Non-HGD exclude the final endoscopy, p53 excludes NA samples') +
  plot.theme + theme(text = element_text(size=16, face='bold'), legend.position = 'bottom')

Same plot, but mapped to the risk categories instead.

```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=9}
back$Pathology = as.character(back$Pathology)
back$Pathology[back$Pathology %in% c('HGD','IMC')] = 'HGD/IMC'
back$Pathology = factor(back$Pathology, levels=c('BE', 'ID','LGD','HGD/IMC'), ordered = T)

riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
                  cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

@@

@@

ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('red','orange2','green3')) + 
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
```

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
# All samples
all.predictions = sapply(pg.samp, function(pt) {
  if (unique(pt$Status) == 'NP') {
    as.integer(pt$Prediction < cutoff)
  } else {
    as.integer(pt$Prediction > cutoff)
  }
})

all.predictions = do.call(rbind, lapply(all.predictions, function(x) {
  cbind.data.frame('correctly.predicted'=sum(x), 'total.samples'=length(x))
}))

all.predictions$Status = ifelse (rownames(all.predictions) %in% subset(sum.patient.data, Status == 'P')$Hospital.Research.ID, 'P', 'NP')

rownames(all.predictions) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(all.predictions))$Patient

all.predictions$pred.ratio = with(all.predictions, correctly.predicted/total.samples)
all.predictions$Patient = rownames(all.predictions)
all.predictions = transform(all.predictions, Patient=reorder(Patient, pred.ratio) )

p = subset(all.predictions, Status == 'P')
np = subset(all.predictions, Status == 'NP')

m = melt(all.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))

>>>>

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')

data = '~/Data/Ellie'
tile.w=c(5e06, 'arms')
#tile.w=5e+06
#tile.w=1e7
#tile.w='arms'

@@

@@

@@

cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')

@@

@@

@@

globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)

@@

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}

@@

write.table(ch, quote=F, sep='\t', row.names = F, col.names = T, file=paste( paste(tile.w, collapse='_')  ,'_features.txt'))
```

@@

#oddPt = names(pg.samp)[which.max(performance.at.1se)]
#subset(sum.patient.data, Hospital.Research.ID == oddPt)

#pg.samp[[oddPt]]$Prediction

# Not the pt with the most samples or endos
#sum.patient.data[which.max(sum.patient.data$total.samples),]
```

```{r, eval=F, echo=F, message=F}
image(cov(dysplasia.df[names(which(labels == 1)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$Prediction))

@@

@@

grid.arrange(roc.plot(roc50,'Probability threshold = 0.5'), roc.plot(roc, 'Probability threshold calculated by ROC'), ncol=2)

roc.plot(roc, 'Copy number model')

@@

```{r p53roc, echo=F, fig.height=2, fig.width=2}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')
```

@@

@@

grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)'), ncol=2)
```

@@

```{r echo=F, fig.height=6, fig.width=8}
get.roc.stat<-function(roc) {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )

@@

dodge = position_dodge(width=0.9)
ggplot(melt(m, measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=0, vjust=-1 ), position=dodge, color='white')  +
  scale_fill_manual(values=c('purple3','dodgerblue3','firebrick3'), name='') + plot.theme + 
  theme(legend.position = 'bottom') + labs(x='', y='Percent', title='ROC Comparison') 
```

@@

## Prediction probability distribution

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
  )

@@

  tb = with(pt, table(Endoscopy.Year, PID))
  tb[tb>0] = 1
  for (yr in names(which(rowSums(tb) > 1))) {
      tmp = pt[which(pt$Endoscopy.Year == yr),]
      tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))

@@

  tb = table(pt[,c('months.before.final', 'PID')])
  for (i in 1:nrow(tb)) {
    pids = tb[i,]
    yrs = which(pids > 0)
    if (length(yrs) > 1) {
    for (j in 2:length(yrs)) 
      pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
    }
  }
  arrange(pt, -Endoscopy.Year, PID)
}

@@

sq = c(0, seq(1,132,12),max(back$months.before.final))
back$brks = cut(back$months.before.final, 
                  breaks=sq, 
                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
                  include.lowest = T, ordered_result = T)
back$brks = factor(back$brks, rev(levels(back$brks)))

The labels in each decile are the ratio of progessor samples that have probabilities within that region. The deciles that are greyed out show insigificant enrichment for the NP/P ratio.

```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
cuts = seq(0,1,0.1)
#cuts = quantile(back$Prediction, probs=seq(0,1,1/5))

@@

qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction), 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))

ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')

pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

@@

annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)

#annotation.table = as.data.frame(annotation.table)
#rownames(annotation.table) = annotation.table$Risk
#annotation.table$Risk = NULL

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = c('green3','orange2','red3'), col=NA, alpha=0.2),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="black", fontface=4L)),
  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
  padding=unit(c(5,5),'mm'))

grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )

riskCols = c('red3','orange','green3')
grid.arrange(
  ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
    scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
    #geom_vline(xintercept=pred.confidence$r2, linetype='longdash', color='grey39') +
    #geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=75,label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + scale_color_manual(values=c('grey','black')) + 
    geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
    geom_text(data=annotation.table %>% group_by(Risk, min,max) %>% summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability')

  ,tableGrob(annotation.table, rows=NULL, theme=tt3),
  nrow=2, heights=c(8,2)
  )

back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
back = back %>% group_by(Risk) %>% mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )

ggplot(back, aes(1:nrow(back), Prediction, color=Risk)) + 
  geom_point(aes(color=Risk, shape=Status)) + geom_errorbar(aes(ymin=Prediction-`se(Risk)`, ymax=Prediction+`se(Risk)`)) +
  geom_hline(data=annotation.table, aes(yintercept=max), color='grey') + plot.theme +
  labs(x='sample', y='model probability', title='Per-sample probabilities')
```

@@

Could just as easily get a jacknife CI per patient

```{r jacknife, echo=T, warning=F, message=F, fig.height=5, fig.width=6}

coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
for (pt in names(pg.samp)) {
  lcf = loo.coefs[[pt]][,1]
  names(lcf) = rownames( loo.coefs[[pt]] )
  sharedcfs = intersect(names(lcf), ch$label)
  coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
}
coefs.per.pt[is.na(coefs.per.pt)] = 0

@@

# per patient
# jack.se = apply(coefs.per.pt, 2, function(x) {
#   jk = jackknife(x, var)
#   jk$jack.se
# })
# js = cbind.data.frame(performance.at.1se, jack.se)
# ggplot(js, aes(rownames(js), performance.at.1se, color=(performance.at.1se > mean(performance.at.1se)+sd(performance.at.1se) | performance.at.1se < mean(performance.at.1se)-sd(performance.at.1se)  ))) + ylim(0.6,1) + geom_point() + 
#   geom_errorbar(aes(ymin=performance.at.1se-jack.se, ymax=performance.at.1se+jack.se)) + theme(legend.position = 'none', axis.text.x = element_text(angle=90, hjust=1))

# in this case jack.se is the SE of the variance 
# per coefficient
ch$jack.se = jack =  apply(coefs.per.pt, 1, function(x) {
  jk = jackknife(x, var)
  jk$jack.se
})
ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
  geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
  coord_flip() +
  plot.theme + labs(x='', title='Jackknifed coefficient CI')

@@

Effectively it looks to me like my coefficients have fairly tight CI if the SE of the jackknife variance is used.

@@

```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=9}
riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
                  cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

@@

#riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
#riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)

@@

ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('red2','orange','green3')) + 
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')

ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=Risk)) + geom_bar(stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('red2','green3')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
all.preds = back %>% group_by(Status, Patient) %>% summarise( 'Low'=sum(length(which(Risk == "Low"))),
                                          'Moderate'=sum(length(which(Risk == "Moderate"))),
                                          'High'=sum(length(which(Risk == "High"))),
                                          'n.samples'=length(Patient) )

all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))
head(all.preds)
ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
  geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=c('green3','orange2','red')) + 
  geom_text(aes(y=1.01, label=n.samples)) +
  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
  plot.theme
